---
title: " Identifying the “scaring away” pattern of wild boars from GPS trajectories in the Swiss Plateau"
author: "Ricarda Ferrari, Michael Niederberger"
output: html_document
---

```{r setup, include=FALSE}
#install.packages("devtools")
#install.packages("trajr")

library(readr)
library(sf)
library(terra)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tmap)
library(sp)
library(devtools)
library(som)
library(trajr)
library (lme4)
library(ComputationalMovementAnalysisData)

devtools::install_github("ComputationalMovementAnalysis/ComputationalMovementAnalysisData")

```


# show data
```{r, message=FALSE, warning=FALSE}
head(wildschwein_BE)
head(wildschwein_metadata)
head(wildschwein_overlap_temp)
head(schreck_agenda)
head(schreck_locations)
```

# Schreck prepare
```{r, message=FALSE, warning=FALSE}
# Convert CRS
schreck_locations_ch <- schreck_locations %>% st_as_sf(coords = c("lon", "lat"), crs = CRS("+init=epsg:4326"), remove = FALSE) 
schreck_locations_ch <- schreck_locations_ch %>% st_transform(crs = 2056)
# Filter scare device regions
schreck_locations_ch <- schreck_locations_ch %>% filter(lat < 47.2 & lon < 7.5)
coordsne <- unlist(st_geometry(schreck_locations_ch)) %>% matrix(ncol=2,byrow=TRUE) %>% as_tibble() %>% setNames(c("N","E"))
schreck_locations_ch$N <- coordsne$E
schreck_locations_ch$E <- coordsne$N
#join 
schreck_locations_ch <- schreck_locations_ch %>% left_join(schreck_agenda, by=c("id"="id"))
schreck_locations_ch$wid <- c(1:25)
schreck_locations_ch <- schreck_locations_ch %>% mutate(wid=as.character(wid))
```


# Data exploratio. Get common data, show examples
```{r, message=FALSE, warning=FALSE}
# Get common samples
sabi <- wildschwein_BE %>% filter(TierName=="Sabine")
# Filter night data
sabi <- sabi %>% filter(day != "Tag" & !is.na(day)) # only night gps
s1 <- schreck_locations_ch[5,]
sabi <- sabi %>% filter(DatetimeUTC > first(s1$datum_on) & DatetimeUTC < first(s1$datum_off))
sabi <- sabi %>% mutate(dist = sqrt((first(s1$N)-N)^2+(first(s1$E)-E)^2))
sabi <- sabi %>% filter(dist < 400)
sabi <- sabi %>% mutate(trip = ifelse(hour(DatetimeUTC) > 16, day(DatetimeUTC)+1, day(DatetimeUTC)))

ggplot() +
  geom_path(data=sabi %>%  filter(trip<11), aes(x=E, y=N, color=factor(trip))) +
  geom_point(data = s1, aes(x=E, y=N), colour="black", size=2) 

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


# Adding closest active schreck (+ distance) to each observation
```{r, message=FALSE, warning=FALSE}
### data
w<-wildschwein_BE %>% mutate(date = as.Date(DatetimeUTC), 
                             time= format(DatetimeUTC, format = "%H:%M:%S"),
                             hour= as.integer(format(DatetimeUTC, format = "%H"))+
                               as.integer(format(DatetimeUTC, format = "%M"))/60)
sl<-schreck_locations

###
s<-data.frame(schreck_locations_ch) ## I had to transform it to a data frame, otherwise something was weird about the coordinates (geometry)
###

s1<-s[!duplicated(s$id),c("id","N","E")] # list with all Schrecks and their location

# Set id to w
w <- w %>% mutate(id = row_number())

## reduce data to certain time frame
w1<-w %>% 
  filter (day!="Tag"&!is.na(day) &
            ((DatetimeUTC > as.Date("2014-05-01") & DatetimeUTC < as.Date("2014-07-04")) | 
                (DatetimeUTC > as.Date("2015-05-20") & DatetimeUTC < as.Date("2015-07-01")) |
                 (DatetimeUTC > as.Date("2016-04-04") & DatetimeUTC < as.Date("2016-10-04")) |
                 (DatetimeUTC > as.Date("2017-04-26") & DatetimeUTC < as.Date("2017-11-18"))))
nrow(w1)


#for(j in 1:nrow(w1)){
### look only at Schrecks that were active on that day
 # s_on<-s[s$datum_on < w1[j,]$DatetimeUTC & s$datum_off > w1[j,]$DatetimeUTC,]
  #if(nrow(s_on)==0)
   # {w1[j,"closest_schreck"]<-"no_Schreck_on"}  
  #else{     
    #for(i in 1:nrow(s_on)){ ## calculate difference between current observation (j) and each schreck location
      #s_on[i,"diff"]<-sqrt((w1[j,"N"]-s_on[i,"N"])^2+(w1[j,"E"]-s_on[i,"E"])^2)} ## add difference of current observation to location into file
#### look at the distance of the closest Schreck, only use it if less than 400m
  #if(min(s_on$diff)>400){
      #w1[j,"closest_schreck"]<-"no_Schreck_witin_400m"} 
  #else{
      #w1[j,"closest_schreck"]<-s_on[s_on$diff==min(s_on$diff),"id"]  ## add closest schreck to each wild boar location
      #w1[j,"distance_to_closest_schreck"]<-s_on[s_on$diff==min(s_on$diff),"diff"]
#}}}

# Save data.frame to spare time
#write.csv(w1, "wildboar_loop.csv")

w1 <- read_delim("wildboar_loop.csv",",")
w1 <- w1 %>% mutate(hour= as.integer(format(DatetimeUTC, format = "%H")),
                tripdate= ifelse(hour < 12, as.Date(DatetimeUTC)-1, as.Date(DatetimeUTC)),
                tripdate2 = as.Date(tripdate, origin="1970-01-01"))

# Merge close wildschweinschreck gps  data to origin wildboar
w <- w %>% left_join(w1 %>% dplyr::select(id, closest_schreck, distance_to_closest_schreck), by="id")
w <- w %>% mutate(closest_schreck= 
  ifelse(is.na(closest_schreck), "no_Schreck_witin_400m", closest_schreck))

#add tripdate
w <- w %>% mutate(hour= as.integer(format(DatetimeUTC, format = "%H")),
                tripdate= ifelse(hour < 12, as.Date(DatetimeUTC)-1, as.Date(DatetimeUTC)),
                tripdate2 = as.Date(tripdate, origin="1970-01-01"))


w %>% dplyr::select(DatetimeUTC, tripdate2)
```

# Separation of the wildboar trips
```{r, message=FALSE, warning=FALSE}

# split time and day in separate columns:
w1$Date <- as.Date(w1$DatetimeUTC)
## order data frame by animal and time
w1<-w1[order(w1$TierName) & order(w1$DatetimeUTC),]


##### Create trips: with shorest distance to schreck (for each tripdate) ########################################

# split time and day in separate columns:
w1$Date <- as.Date(w1$DatetimeUTC)

## order data frame by animal and time
w1<-w1[order(w1$TierName) & order(w1$DatetimeUTC),]

## empty column for trip id
w$tripID_dist<-"NA"
w$start_dist<-"no"
w$isNearestPoint <- "no"
animals<-unique(w1$TierName)

for(j in 1:length(animals)){
  days<-unique(w1[w1$TierName==animals[j] & !is.na(w1$TierName),]$tripdate2) 
  if(length(days)!=0){
    for(i in 1:length(days)){
      n<-w1[w1$TierName==animals[j] & w1$tripdate2==days[i],]
      if(nrow(n[!is.na(n$distance_to_closest_schreck),])==0){nmin<-NA} else{
      nmin<-min(n$distance_to_closest_schreck,na.rm=T)}
      if(!is.na(nmin)){ ## only continue if nmin is not NA
        if(nmin<=400){ ## only continue if minimal distance to schreck is less than 400m
      n<-n[n$distance_to_closest_schreck==nmin & !is.na(n$distance_to_closest_schreck),]
      w[w$TierName==animals[j] & w$DatetimeUTC == n$DatetimeUTC, "isNearestPoint"] <- "yes"
        w[w$TierName==animals[j] & (w$DatetimeUTC >= n$DatetimeUTC-4*60*60) & (w$DatetimeUTC <= n$DatetimeUTC+4*60*60),"tripID_dist"]<-paste(animals[j],i, sep="_")                               
        w[w$TierName==animals[j] & w$DatetimeUTC==n$DatetimeUTC,"start_dist"]<-"yes"
      }}}}}

# Only trips in the night
w <- w %>% group_by(TierID, tripdate2) %>% mutate(isTrip= max(tripID_dist) != 'NA',
                                          tripIDnight = ifelse((hour < 9.1 | hour >= 17.9)&isTrip,
                                                              max(tripID_dist), NA)) %>% ungroup()


# length(unique(w$tripIDnight)); table(w[!is.na(w$tripIDnight),]$isNearestPoint)

#List with closest_distance to schreck for each trip
trip_list <- w %>% 
  filter (isNearestPoint=="yes" & !is.na (tripIDnight))

nrow(trip_list)

head(data.frame(trip_list))
```

# Overview data
```{r, message=FALSE, warning=FALSE}
### plot
pa<-ggplot(w,aes(x=date,y=TierName))+geom_line()
pa<-pa+ylab("animal ID")+scale_x_date(breaks = "3 month", minor_breaks = "1 month")+xlab("date")+theme_bw()
pa
#ps<-ggplot(s[s$id %in% w$closest_schreck,],aes(y=id,yend=id,x=datum_on,xend=datum_off))+geom_segment()
#ps

```

# Wildschweinschreck
```{r, message=FALSE, warning=FALSE}

# Determine mean speaker orientation
schreck_locations_ch <- schreck_locations_ch %>% 
  mutate(ausrichtung_mean = (ausrichtung_max - ausrichtung_min) / 2 +ausrichtung_min,
         ausrichtung_dir = ifelse(ausrichtung_mean >= 45 & ausrichtung_mean < 135, "E", 
                                  ifelse(ausrichtung_mean > 135 & ausrichtung_mean < 225, "S",
                                         ifelse(ausrichtung_mean > 225 & ausrichtung_mean < 315, "W", "N")))) 


# Show example of orientation of wildboar scare device with loudness and direction
s1 <- schreck_locations_ch[5,]

wildboar_closeup <- wildschwein_BE %>% 
  filter(TierName=="Sabine" & DatetimeUTC > as_datetime('2015-06-08 18:30:43') & DatetimeUTC < as_datetime('2015-06-09 10:30:43')) %>% 
  mutate(dist = sqrt((first(s1$N)-N)^2+(first(s1$E)-E)^2)) %>% mutate(triptime0 = as.numeric((DatetimeUTC - min(DatetimeUTC))) / 60) %>% arrange(.,triptime0)

schreck_orientation <- s1 %>% mutate(length=lautstaerke*lautstaerke/100)
rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}

schreck_orientation$asurichtung_mean <- (schreck_orientation$ausrichtung_max-
                                           schreck_orientation$ausrichtung_min) / 2 +
                                        schreck_orientation$ausrichtung_min
schreck_orientation$ausrichung_meanE = schreck_orientation$E[1] + schreck_orientation$length[1] * 2*
  cos(deg2rad(360-270-(schreck_orientation$asurichtung_mean[1])))
schreck_orientation$ausrichung_meanN = schreck_orientation$N[1] + schreck_orientation$length[1] *2* 
  sin(deg2rad(360-270-(schreck_orientation$asurichtung_mean[1])))

schreck_orientation$ausrichung_minE = schreck_orientation$E[1] + schreck_orientation$length[1] * 
  cos(deg2rad(360-270-schreck_orientation$ausrichtung_min[1]))
schreck_orientation$ausrichung_minN = schreck_orientation$N[1] + schreck_orientation$length[1] * 
  sin(deg2rad(360-270-schreck_orientation$ausrichtung_min[1]))
schreck_orientation$ausrichung_maxE = schreck_orientation$E[1] + schreck_orientation$length[1] * 
  cos(deg2rad(360-270-schreck_orientation$ausrichtung_max[1]))
schreck_orientation$ausrichung_maxN = schreck_orientation$N[1] + schreck_orientation$length[1] * 
  sin(deg2rad(360-270-schreck_orientation$ausrichtung_max[1]))

x_coord <- c(schreck_orientation$E[1],  schreck_orientation$ausrichung_minE[1],
             schreck_orientation$ausrichung_meanE[1],
             schreck_orientation$ausrichung_maxE[1], schreck_orientation$E[1])
y_coord <- c(schreck_orientation$N[1], schreck_orientation$ausrichung_minN[1],
             schreck_orientation$ausrichung_meanN[1],
             schreck_orientation$ausrichung_maxN[1], schreck_orientation$N[1])

poly <- st_polygon(list(matrix(c(x_coord, y_coord),ncol=2, byrow=FALSE))) 
pos <- data.frame(x=x_coord, y=y_coord, id=c(1,2,3,4,5))

# Show example of scare device orientation and loudness
ggplot() +
  geom_path(data=wildboar_closeup, aes(x=E, y=N, color=triptime0)) +
  geom_point(data = s1, aes(x=E, y=N), colour="black", size=2) +
  geom_point(data = schreck_orientation, aes(x=ausrichung_meanE, y=ausrichung_meanN), 
             colour="brown", size=2) +
  geom_polygon(data=pos, aes(x=x, y = y), fill="orange", alpha=0.4)
```

# Wilddboar scare effect properties
```{r, message=FALSE, warning=FALSE}
# Get all trips
trips <- w %>% filter(!is.na(tripIDnight)) %>% group_by(tripIDnight)
trips 

wildboar_trip_scared <- data.frame(matrix(ncol = 22, nrow = 0))
x <- c("id", "TierName","triptime0", "approachingRate", "approachingRateRelative",
       "approachingRateAbsolute", "speed", "sinousity", "linedist", "acceleration", "speedDiff3", 
       "dist","E", "N","x", "y", "directionRelative",
       "closest_schreck", "tripIDnight", "DatetimeUTC", "hour", "day")
colnames(wildboar_trip_scared) <- x
# Amount of trips
unique(trips$tripIDnight)

for (u in 1:length(unique(trips$tripIDnight)))
{
  wildboar_trip <- trips %>% filter(tripIDnight==unique(trips$tripIDnight)[u])
  
  # Only take full trips
  if (length(wildboar_trip$TierID) > 61)
  {
    # Get wildboard schreck associated to wild boar
    schreck <- schreck_locations_ch %>%  filter(id == max(wildboar_trip$closest_schreck))
    # Calculate distance and time differences
    wildboar_trip <- wildboar_trip %>% 
      mutate(dist = sqrt((first(schreck$N)-N)^2+(first(schreck$E)-E)^2),
        triptime0 = as.numeric((DatetimeUTC - min(DatetimeUTC))) / 60,
        triptimeDiff = (triptime0-lag(triptime0))*60,
        distanceAbsolute = ((E- lag(E))^2 + (N-lag(N))^2)^0.5) %>% arrange(.,triptime0)
    
    # Calculate approaching rate
    wildboar_trip <- wildboar_trip %>%  
      mutate(approachingRate = lag(dist)-dist,
             approachingRateAbsolute = approachingRate / (triptimeDiff),
             approachingRateRelative = approachingRate / (distanceAbsolute),
             approachingRateRelative = ifelse(is.na(approachingRateRelative), 0, approachingRateRelative))
    # Calculate speed & acceleration
    wildboar_trip <- wildboar_trip %>%
      mutate(speed = round(distanceAbsolute / triptimeDiff, 4),
             acceleration = round((lead(speed)-speed)  * 60 / lead(triptimeDiff), 4),
             speedDiff3 = round(((speed+lead(speed)+lead(speed, 2))/3 - 
                                      (lag(speed)+lag(speed, 2)+lag(speed, 3)/3)), 4))

    # Relative coordinates (trajectory)
    wildboar_trip <- wildboar_trip %>% mutate(x=E-first(schreck$E), y=N-first(schreck$N))
    coords <- data.frame(x = wildboar_trip$x, 
                         y = wildboar_trip$y, 
                         times = wildboar_trip$triptime0)
    # Set all sinousitiy to 0
    wildboar_trip$sinousity = replicate(length(coords$x), NA)
    for (i in 1:(length(coords$x)-3)) {
      # Create a trajectory from the coordinates
      trj <- TrajFromCoords(coords[(i):(i+3),])
      # Rescale stepsize
      trj_re <- TrajRediscretize(trj, 1)
      # Calculate sinousity
      wildboar_trip$sinousity[i] = TrajSinuosity(trj_re, compass.direction = TRUE) %>% 
        round(digits = 4)
    }
    
    # Get proximity of movement
    wildboar_trip <- wildboar_trip %>%
     mutate(linedist = ((x^2+y^2)^0.5 + ((x-lag(x))^2+(y-lag(y))^2)^0.5)/2)
    # Get relative direction to scare device
    wildboar_trip <- wildboar_trip %>% mutate(directionRelative=
                                                ifelse(abs(x) > abs(y),
                                                       ifelse(x>0, "E", "W"),
                                                       ifelse(y>0, "N", "S")))
    
    # Select
    wildboar_trip_scared <- wildboar_trip_scared %>% 
      rbind(wildboar_trip %>% 
              dplyr::select(id, TierName, triptime0, approachingRate, approachingRateRelative,
                            approachingRateAbsolute, speed, sinousity, linedist, acceleration, 
                            speedDiff3, dist, E, N, x, y, directionRelative, closest_schreck, tripIDnight, 
                            DatetimeUTC, hour, day))
  }
}
head(wildboar_trip_scared)
# Scaled values, calculate scare effect and set "scaring away" points
wildboar_trip_scared <- wildboar_trip_scared %>% group_by(tripIDnight) %>% 
  mutate(
    approachingRateRelativeS = (approachingRateRelative-min(approachingRateRelative, na.rm = TRUE)) /
      (max(approachingRateRelative, na.rm = TRUE)- min(approachingRateRelative, na.rm = TRUE)),
    accelerationS = (acceleration-min(acceleration, na.rm = TRUE))/
                     (max(acceleration, na.rm = TRUE) - min(acceleration, na.rm = TRUE)),
    sinousityS = (sinousity-min(sinousity, na.rm = TRUE))/
                     (max(sinousity, na.rm = TRUE)- min(sinousity, na.rm = TRUE)),
    distanceS = (dist-min(dist, na.rm = TRUE)) / (max(dist, na.rm = TRUE) - min(dist, na.rm = TRUE)),
    speedDiff3S = (speedDiff3-min(speedDiff3, na.rm=TRUE)) / 
      (max(speedDiff3, na.rm=TRUE)-min(speedDiff3, na.rm = TRUE)),
    approachingRateRelativeSlead = lead(approachingRateRelativeS),
    scareEffect = accelerationS - sinousityS - distanceS - lead(approachingRateRelativeS) + speedDiff3S,
    scareEffectShow = ifelse(scareEffect < 0, -0.1, scareEffect),
    scary= scareEffect > 1.4,
    scary= ifelse(is.na(scary), FALSE, scary),
    scary= ifelse(scary&lag(scary), FALSE, scary)) %>% 
  ungroup()

# Get scared points
wildboar_scared <- wildboar_trip_scared %>% filter(scary)

# Get scared trips
wildboar_trip_scared_True <- wildboar_trip_scared %>% 
  group_by(tripIDnight) %>% 
  mutate(tripScared = sum(scary, na.rm = TRUE)) %>% 
  filter(tripScared > 0)

# Join trips with scared wild boars
w <- w %>% left_join(wildboar_trip_scared_True %>% ungroup() %>% 
                       dplyr::select(id, scareEffect, triptime0, accelerationS,
                                     approachingRateRelativeSlead, speedDiff3S, sinousityS,
                                     distanceS, x, y, scary), by="id")
```

# Plot all found shrecks
```{r, message=FALSE, warning=FALSE}
# Amount of scared trips
unique(wildboar_trip_scared$tripIDnight)
unique(wildboar_trip_scared_True$tripIDnight)
length(unique(wildboar_trip_scared_True$tripIDnight))

for (u in 1:length(unique(wildboar_trip_scared_True$tripIDnight))) 
{
  test <- wildboar_trip_scared_True %>% filter(tripIDnight==unique(wildboar_trip_scared_True$tripIDnight)[u])
  print(first(test$DatetimeUTC))
  print(first(test$tripIDnight))
  # Variable values
  print(ggplot(data=test) + 
    geom_bar(stat="identity", aes(x=triptime0/60, y=scareEffectShow),fill = "grey", size=1.3)+ 
    geom_line( aes(x=triptime0/60, y=distanceS, alpha=0.8), color="blue", size=1.3) +
    geom_line( aes(x=triptime0/60, y=approachingRateRelativeSlead, alpha=0.8), color="green", size=1.3) +
    geom_line( aes(x=triptime0/60, y=speedDiff3S, alpha=0.8), color="red", size=1.3) +
    geom_line( aes(x=triptime0/60, y=sinousityS, alpha=0.8), color="turquoise", size=1.3) +
    geom_line( aes(x=triptime0/60, y=accelerationS, alpha=0.8), color="orange", size=1.3) +
    ggtitle(paste("Wildboar trip", first(test$tripIDnight), "started at", first(test$DatetimeUTC))) +
    xlab("Hour passed by from 18:00 p.m.") + ylab("Scaled variables (0 to 1) and scare effect (-3 to 2)"))
  
    testClose <- test %>% 
      filter(scary | lead(scary) | lead(scary,2)| lead(scary,3)| lead(scary,4) | lag(scary,1)| lag(scary,2)| lag(scary,3)| lag(scary,4))
  print(first(test$DatetimeUTC))
  print(first(test$tripIDnight))
  print(ggplot(data=testClose) + 
    geom_bar(stat="identity", aes(x=(triptime0-testClose$triptime0[5]), y=scareEffectShow),
             fill = "grey", size=1.3)+
    geom_line(aes(x=(triptime0-testClose$triptime0[5]), y=distanceS, alpha=0.8, colour="Distance"),
               color="blue", size=1.3) +
    geom_line( aes(x=(triptime0-testClose$triptime0[5]), 
                   y=approachingRateRelativeSlead, alpha=0.8), color="green", size=1.3) +
    geom_line( aes(x=(triptime0-testClose$triptime0[5]), y=accelerationS, alpha=0.8), color="orange", size=1.3) +
    geom_line( aes(x=(triptime0-testClose$triptime0[5]), y=speedDiff3S, alpha=0.8), color="red", size=1.3) +
    geom_line( aes(x=(triptime0-testClose$triptime0[5]), y=sinousityS, alpha=0.8), color="turquoise", size=1.3) +
    ggtitle(paste("Wildboar trip", first(testClose$tripIDnight), "1 hour before and after scare effect at",
                  testClose$DatetimeUTC[5])) +
    xlab("Minutes before or after the scare effect") + 
    ylab("Scaled variables (0 to 1) and scare effect (-3 to 2)")+ 
    theme(legend.position = "bottom"))
  
  print(ggplot() +
    geom_path(data= testClose, aes(x=E, y=N, color=(triptime0-testClose$triptime0[5]))) +
    geom_point(data = schreck_locations_ch %>% filter(id == max(test$closest_schreck)), 
               aes(x=E, y=N), colour="black", size=2) +
    geom_point(data = test %>% filter(scary), aes(x=E, y=N), color="red")+
    ggtitle(paste("Wildboar trip", first(test$tripIDnight), "started at", first(test$DatetimeUTC))))
  
  print(ggplot() +
    geom_path(data= test, aes(x=E, y=N, color=triptime0)) +
    geom_point(data = schreck_locations_ch %>% filter(id == max(test$closest_schreck)), 
               aes(x=E, y=N), colour="black", size=2) +
    geom_point(data = test %>% filter(scary), aes(x=E, y=N), color="red")+
    ggtitle(paste("Wildboar trip", first(testClose$tripIDnight), "1 hour before and after scare effect at",
                  testClose$DatetimeUTC[5])))
}
```

# Calculate noise exposition at scared data points and for trips.
```{r, message=FALSE, warning=FALSE}
# Get scared points and calculate sound volume
wildboar_scared_schreck <- wildboar_scared %>% 
  left_join(schreck_locations_ch %>% rename(schreckID = id), by=c("closest_schreck"="schreckID"))  %>% 
  filter(datum_on < DatetimeUTC & datum_off > DatetimeUTC)

wildboar_scared_schreck <- wildboar_scared_schreck %>% 
  select(id, tripIDnight, TierName, DatetimeUTC, ausrichtung_dir, directionRelative, dist, lautstaerke, scareEffect, modus, day) %>% 
  mutate(sameDirection=ausrichtung_dir==directionRelative,
         volumeDir = ifelse(sameDirection, lautstaerke, 
                           ifelse(ausrichtung_dir=='S'&directionRelative=='N' | 
                                    ausrichtung_dir=='N'& directionRelative=='S' | 
                                    ausrichtung_dir=='W'& directionRelative=='E'| 
                                    ausrichtung_dir=='E'& directionRelative=='W', 
                                  lautstaerke-10, lautstaerke-5)),
         soundDistance = volumeDir - abs(20 * log(1/(1.581*dist), 10)),
         soundDistance = ifelse(day=="Tag", -20, soundDistance),
         schreckProbabily = soundDistance *100/40)

ggplot(data=wildboar_scared_schreck) +
  geom_point(aes(x= scareEffect, y=soundDistance, color=soundDistance>10), size=2)

# Get scared trips and calculate sound volume 
wildboar_trips <- wildboar_trip_scared %>% group_by(tripIDnight) %>% 
  filter(scareEffect == max(scareEffect, na.rm=TRUE)) %>% ungroup() %>% 
  left_join(schreck_locations_ch, by=c("closest_schreck"="id"))  %>% 
  filter(datum_on < DatetimeUTC & datum_off > DatetimeUTC)

wildboar_trips <- wildboar_trips %>% 
  select(id, tripIDnight, TierName, DatetimeUTC, scary, ausrichtung_dir, directionRelative, 
         dist, lautstaerke, scareEffect, modus, day) %>% 
  mutate(sameDirection=ausrichtung_dir==directionRelative,
         volumeDir = ifelse(sameDirection, lautstaerke, 
                           ifelse(ausrichtung_dir=='S' & directionRelative=='N' | 
                                    ausrichtung_dir=='N' & directionRelative=='S' | 
                                    ausrichtung_dir=='W' & directionRelative=='E'| 
                                    ausrichtung_dir=='E' & directionRelative=='W', 
                                  lautstaerke-10, lautstaerke-5)),
         soundDistance = volumeDir - abs(20 * log(1/(1.581*dist), 10)),
         soundDistance = ifelse(day=="Tag", -20, soundDistance),
        scareSound= ifelse(soundDistance < 10, FALSE, scary))

wildboar_trips$sum <- replicate(length(wildboar_trips$TierName), 1)

# Filter scare effect trips by scare sound
wildboar_scared2 <- wildboar_scared %>% 
  left_join(wildboar_trips %>% select(id, soundDistance, scareSound), by=c("id"="id"))
wildboar_scared2 <- wildboar_scared2 %>% filter(scareSound)
wildboar_scared2$ScareAndSound <- replicate(length(wildboar_scared2$TierName), TRUE)
w <- w %>% left_join(wildboar_scared2 %>% select(id, soundDistance, ScareAndSound),by=c('id'='id'))

# Get scared trips and calculate sound volume, select most intense sound point per trip and highest scare effect per trip.
wildboar_trips2 <- wildboar_trip_scared  %>% 
  left_join(schreck_locations_ch, by=c("closest_schreck"="id"))  %>% 
  filter(datum_on < DatetimeUTC & datum_off > DatetimeUTC) %>% 
  select(tripIDnight, TierName, DatetimeUTC, scary, ausrichtung_dir, directionRelative, 
         dist, lautstaerke, scareEffect, modus, day) %>% 
  mutate(sameDirection=ausrichtung_dir==directionRelative,
         volumeDir = ifelse(sameDirection, lautstaerke, 
                           ifelse(ausrichtung_dir=='S' & directionRelative=='N' | 
                                    ausrichtung_dir=='N' & directionRelative=='S' | 
                                    ausrichtung_dir=='W' & directionRelative=='E'| 
                                    ausrichtung_dir=='E' & directionRelative=='W', 
                                  lautstaerke-10, lautstaerke-5)),
         soundDistance = volumeDir - abs(20 * log(1/(1.581*dist), 10)),
         soundDistance = ifelse(day=="Tag", -20, soundDistance),
         schreckProbabily = soundDistance *100/40) 

wildboar_trips2 <- wildboar_trips2 %>% group_by(tripIDnight) %>% 
  mutate(scareEffectMaxTrip=max(scareEffect)) %>% 
  filter(soundDistance == max(soundDistance, na.rm=TRUE)) %>% ungroup() 
wildboar_trips2$sum <- replicate(length(wildboar_trips2$TierName), 1)

ggplot(data=wildboar_trips2) +
  geom_point(aes(x= scareEffectMaxTrip, y=soundDistance, color=soundDistance>10 & scareEffectMaxTrip > 1.4), size=2)


# Show plot of trips (total, noise exposed, scared, and both)
ggplot(data=wildboar_trips) + 
    geom_bar(stat="identity", aes(x=TierName, y=sum),
             fill = "grey", size=1.3) +
    geom_bar(data = wildboar_trips2 %>% filter(soundDistance > 20), stat="identity", aes(x=TierName, y= sum),
             fill = "#ff9999", size=1.3) +
    geom_bar(data = wildboar_trips %>% filter(scary), stat="identity", aes(x=TierName, y=sum),
             fill = "#e83c3c", size=1.3) +
  geom_bar(data = wildboar_trips %>% filter(soundDistance > 10 & scary), stat="identity", aes(x=TierName, y= sum),
             fill = "darkred", size=1.3) +
    xlab("Wild boar animal") + 
    ylab("Amount of trips")

wildboar_trips %>% filter(soundDistance > 20 & scary)
schreck_locations_ch
```


# How long does the scare-off effect last
```{r, message=FALSE, warning=FALSE}
wildboar_trip_scared$return <- replicate(length(wildboar_trip_scared$N), "NA")
wildboar_trip_scared$return <- as.character(wildboar_trip_scared$return)
wildboar_trip_scared$return_hrs <- NA

### looking after end of trip
trip_list$hour<- as.integer(format(trip_list$DatetimeUTC, format = "%H")) ## add hour variable to trip list
trip_list$date2<-trip_list$date
trip_list[trip_list$hour<12,]$date2<-trip_list[trip_list$hour<12,]$date-1

for(j in 1:nrow(trip_list)){
  wildboar_trip_scared$return <- as.character(wildboar_trip_scared$return)
  wildboar_trip_scared$return_hrs <- as.numeric(as.character(wildboar_trip_scared$return_hrs))
  tripend<-paste(trip_list$date2[j]+1,"09:45:00 UTC",sep=" ") ## end of trip
  under400 <- wildboar_trip_scared[wildboar_trip_scared$TierName==trip_list$TierName[j] & wildboar_trip_scared$DatetimeUTC>tripend & wildboar_trip_scared$closest_schreck==trip_list$closest_schreck[j],]  
  under400 <- under400[order(under400$DatetimeUTC),]
  if(nrow(under400)==0){wildboar_trip_scared[wildboar_trip_scared$tripIDnight==trip_list$tripIDnight[j] & !is.na(wildboar_trip_scared$tripIDnight),]$return <- "never"
  wildboar_trip_scared[wildboar_trip_scared$tripIDnight==trip_list$tripIDnight[j] & !is.na(wildboar_trip_scared$tripIDnight), "return_hrs"] <- NA
  } 
  else{
    wildboar_trip_scared[wildboar_trip_scared$tripIDnight==trip_list$tripIDnight[j] & !is.na(wildboar_trip_scared$tripIDnight), "return"] <- as.character(under400[1,]$DatetimeUTC) 
    wildboar_trip_scared$return <- as.POSIXct(wildboar_trip_scared$return, format="%Y-%m-%d %H:%M:%S",tz="UTC")
    return_time <- as.numeric(as.character(difftime(under400[1,]$DatetimeUTC,trip_list[j,]$DatetimeUTC, units="hours")))
    wildboar_trip_scared[wildboar_trip_scared$tripIDnight==trip_list$tripIDnight[j] & !is.na(wildboar_trip_scared$tripIDnight),"return_hrs"] <- return_time
  }}



### create file with one line per trip (the observation with the highest scare away effect)
wildboar_trip_scared$highest<-"no"
triplist <- unique(wildboar_trip_scared$tripIDnight)
for(i in 1:length(triplist)){
  list<-wildboar_trip_scared[wildboar_trip_scared$tripIDnight==triplist[i],"scareEffect"]
  maxscare<-max(list,na.rm=T)
  if(!is.na(maxscare)){
  wildboar_trip_scared[wildboar_trip_scared$scareEffect==maxscare & !is.na(wildboar_trip_scared$scareEffect),"highest"]<-"yes"}}

### unique file
ww<-wildboar_trip_scared[wildboar_trip_scared$highest=="yes",]

### creating new column scared_away (yes if at least one observation is classified as scared for a given trip)
ww$scared_away <- "no"
for(i in 1:length(triplist)){
 list<-ww[ww$tripIDnight==triplist[i],"scary"]
  if(TRUE %in% list){ww[ww$tripIDnight==triplist[i],]$scared_away<-"yes"}}



tt<-ww
head(data.frame(tt))
 ph<-ggplot(tt,aes(x=return_hrs))+geom_histogram(binwidth = 6)+theme_bw()+xlab("hours until animal returned within 400 of schreck [m]")
 ph

 ### plot return hrs for scared versus non-scared:
sp<-ggplot(tt,aes(y=return_hrs,x=scared_away))+geom_boxplot()+theme_bw()+xlab("scared away")+ylab("hours until return")
sp  

#### scared away versus not
  meanreturn1<-aggregate(return_hrs~scared_away, data=tt,mean)
 nreturn1<-aggregate(return_hrs~scared_away, data=tt,length)
 sdreturn1<-aggregate(return_hrs~scared_away, data=tt,sd)
 
## calculate standard error from sample size (n) and standard deviation (sd)
 meanreturn1$SE<-sdreturn1$return_hrs/sqrt(sdreturn1$return_hrs)
 

 ### linear mixed model (accounting for repeated measures from the same individual)
m1<-lmer(return_hrs~scared_away +(1|TierName),data=tt) 
summary(m1)
drop1(m1,test="Chisq")
confint(m1)   
# assumptions of normal distribution of residuals violeted
# Main issue: only seven trips for scare equal TRUE
#### due to non-normality of residuals: non paratmetric test instead (however, not controlled for individual)
wilcox.test(return_hrs~scared_away,data=tt)

#### 
 meanreturn<-aggregate(return_hrs~TierName, data=tt,mean)
 nreturn<-aggregate(return_hrs~TierName, data=tt,length)
 sdreturn<-aggregate(return_hrs~TierName, data=tt,sd)
 
 ## calculate standard error from sample size (n) and standard deviation (sd)
 meanreturn$SE<-sdreturn$return_hrs/sqrt(sdreturn$return_hrs)
 meanreturn$lwr<-meanreturn$return_hrs-meanreturn$SE
 meanreturn$upr<-meanreturn$return_hrs+meanreturn$SE

 pm<-ggplot(meanreturn,aes(y=return_hrs,x=TierName))+geom_point()
 pm<-pm+geom_errorbar(data=data.frame(meanreturn),aes(ymin=lwr,ymax=upr),width=0.4)+theme_bw()+xlab("animal ID")+ylab("mean and SE of hours until return")
 
 
 ## boxplot
 ptt<-ggplot(tt,aes(y=return_hrs,x=TierName))+geom_boxplot()+xlab("animal ID")+theme_bw()
 ptt
 

 ## linear model
 mp<-ggplot(tt,aes(y=return_hrs,x=scareEffect))+geom_point()+geom_smooth(method="lm")+theme_bw()+ylab("hours until return")
 mp
m<-lmer(log(return_hrs)~scareEffect +(1|TierName),data=tt) ## TierName as random effect to control for repeated mesures per animal
qqnorm(resid(m)); qqline(resid(m))
summary(m)     
drop1(m,test="Chisq")
  

### correlation
cor.test(tt$scareEffect,tt$return_hrs, method="spearman")  # spearman because of non-normality of residuals

```
