---
title: "week5-ex"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(sf)
library(terra)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tmap)
library(sp)
library(devtools)

install.packages("devtools")

devtools::install_github("ComputationalMovementAnalysis/ComputationalMovementAnalysisData")
library(ComputationalMovementAnalysisData)
```


# show data
```{r cars}
head(wildschwein_BE)
head(wildschwein_metadata)
head(wildschwein_overlap_temp)
head(schreck_agenda)
head(schreck_locations)
```

# Schreck prepare
```{r }
schreck_locations_ch <- schreck_locations %>% st_as_sf(coords = c("lon", "lat"), crs = CRS("+init=epsg:4326"), remove = FALSE)  #%>% st_transform(crs = 2056)
schreck_locations_ch <- schreck_locations_ch %>% st_transform(crs = 2056)
schreck_locations_ch <- schreck_locations_ch %>% filter(lat < 47.2 & lon < 7.5)
coordsne <- unlist(st_geometry(schreck_locations_ch)) %>% matrix(ncol=2,byrow=TRUE) %>% as_tibble() %>% setNames(c("N","E"))
schreck_locations_ch$N <- coordsne$E
schreck_locations_ch$E <- coordsne$N
#join 
schreck_locations_ch <- schreck_locations_ch %>% left_join(schreck_agenda, by=c("id"="id"))
```

# Get common data, show examples
```{r }
# Get common samples
head(wildschwein_BE)
head(schreck_locations_ch)
sabi <- wildschwein_BE %>% filter(TierName=="Sabine")
# Filter night data
sabi <- sabi %>% filter(day != "Tag" & !is.na(day)) # only night gps

s1 <- schreck_locations_ch[5,]
sabi <- sabi %>% filter(DatetimeUTC > first(s1$datum_on) & DatetimeUTC < first(s1$datum_off))
sabi <- sabi %>% mutate(dist = sqrt((first(s1$N)-N)^2+(first(s1$E)-E)^2))
sabi <- sabi %>% filter(dist < 400)
sabi <- sabi %>% mutate(trip = ifelse(hour(DatetimeUTC) > 16, day(DatetimeUTC)+1, day(DatetimeUTC)))


ggplot() +
  geom_path(data=sabi %>%  filter(trip<11), aes(x=E, y=N, color=factor(trip))) +
  geom_point(data = s1, aes(x=E, y=N), colour="black", size=2) 

# Get example 1
wildboar_closeup <- wildschwein_BE %>% 
  filter(TierName=="Sabine" & DatetimeUTC > as_datetime('2015-06-09 22:30:43') & DatetimeUTC < as_datetime('2015-06-10 14:30:43')) %>% 
  mutate(dist = sqrt((first(s1$N)-N)^2+(first(s1$E)-E)^2)) %>% mutate(time0 = as.numeric((DatetimeUTC - min(DatetimeUTC))) / 60)
ggplot() +
  geom_path(data=wildboar_closeup, aes(x=E, y=N, color=time0)) +
  geom_point(data = s1, aes(x=E, y=N), colour="black", size=2) 
# Get example 2
wildboar_closeup1 <- wildschwein_BE %>% 
  filter(TierName=="Sabine" & DatetimeUTC > as_datetime('2015-06-08 18:30:43') & DatetimeUTC < as_datetime('2015-06-09 14:30:43')) %>% 
  mutate(dist = sqrt((first(s1$N)-N)^2+(first(s1$E)-E)^2)) %>% mutate(time0 = as.numeric((DatetimeUTC - min(DatetimeUTC))) / 60) %>% arrange(.,time0)
ggplot() +
  geom_path(data=wildboar_closeup1, aes(x=E, y=N, color=time0)) +
  geom_point(data = s1, aes(x=E, y=N), colour="black", size=2) 
# Get example 2
wildboar_closeup2 <- wildschwein_BE %>% 
  filter(TierName=="Sabine" & DatetimeUTC > as_datetime('2015-06-07 18:30:43') & DatetimeUTC < as_datetime('2015-06-08 14:30:43')) %>% 
  mutate(dist = sqrt((first(s1$N)-N)^2+(first(s1$E)-E)^2)) %>% mutate(time0 = as.numeric((DatetimeUTC - min(DatetimeUTC))) / 60) %>% arrange(.,time0)
ggplot() +
  geom_path(data=wildboar_closeup2, aes(x=E, y=N, color=time0)) +
  geom_point(data = s1, aes(x=E, y=N), colour="black", size=2) 

# Plot
head(schreck_locations_ch)
ggplot() +
  geom_sf(data = schreck_locations_ch, color='black') +
  geom_sf(data=wildschwein_BE %>% filter(TierName=="Sabi"), color="blue")

ggplot() +
  geom_point(data = schreck_locations_ch, aes(x=E, y=N, color='red')) +
  geom_point(data=sabi, aes(x=E, y=N,color="blue")) + ylim(1200000, 1210000) +  xlim(2568000, 2578000)
sabi %>% filter(day==10)



```

```{r}
# calculate properties of scarce (approaching rate, speed)
wildboar_closeup <- wildschwein_BE %>% 
  filter(TierName=="Sabine" & DatetimeUTC > as_datetime('2015-06-09 22:30:43') & DatetimeUTC < as_datetime('2015-06-10 14:30:43')) %>% 
  mutate(dist = sqrt((first(s1$N)-N)^2+(first(s1$E)-E)^2)) %>% mutate(time0 = as.numeric((DatetimeUTC - min(DatetimeUTC))) / 60)
schreck_orientation <- s1 %>% mutate(length=lautstaerke*lautstaerke/100)
rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}

schreck_orientation$asurichtung_mean <- (schreck_orientation$ausrichtung_max-
                                           schreck_orientation$ausrichtung_min) / 2 +
                                        schreck_orientation$ausrichtung_min
schreck_orientation$ausrichung_meanE = schreck_orientation$E[1] + schreck_orientation$length[1] * 2*
  cos(deg2rad(270-(360-schreck_orientation$asurichtung_mean[1])))
schreck_orientation$ausrichung_meanN = schreck_orientation$N[1] + schreck_orientation$length[1] *2* 
  sin(deg2rad(270-(360-schreck_orientation$asurichtung_mean[1])))

#Dirinst = atan2(-Uinst,-Vinst) * DperR
#Dirgeo = Dirinst + Vaz

schreck_orientation$ausrichung_minE = schreck_orientation$E[1] + schreck_orientation$length[1] * 
  cos(deg2rad(270-(360-schreck_orientation$ausrichtung_min[1])))
schreck_orientation$ausrichung_minN = schreck_orientation$N[1] + schreck_orientation$length[1] * 
  sin(deg2rad(270-(360-schreck_orientation$ausrichtung_min[1])))
schreck_orientation$ausrichung_maxE = schreck_orientation$E[1] + schreck_orientation$length[1] * 
  cos(deg2rad(270-(360-schreck_orientation$ausrichtung_max[1])))
schreck_orientation$ausrichung_maxN = schreck_orientation$N[1] + schreck_orientation$length[1] * 
  sin(deg2rad(270-(360-schreck_orientation$ausrichtung_max[1])))


x_coord <- c(schreck_orientation$E[1],  schreck_orientation$ausrichung_minE[1],
             schreck_orientation$ausrichung_meanE[1],
             schreck_orientation$ausrichung_maxE[1], schreck_orientation$E[1])
y_coord <- c(schreck_orientation$N[1], schreck_orientation$ausrichung_minN[1],
             schreck_orientation$ausrichung_meanN[1],
             schreck_orientation$ausrichung_maxN[1], schreck_orientation$N[1])
p = Polygon(cbind(x_coord, y_coord))
ps = Polygons(list(p),1)
sps = SpatialPolygons(list(ps))
plot(sps)
#schreck_orientation$polygon[1] = sps[1]
poly <- st_polygon(list(matrix(c(x_coord, y_coord),ncol=2, byrow=FALSE))) %>% st_geometry(poly) %>% st_set_crs(2056)
schreck_orientation$polygon <- poly 
                                                                        

ggplot() +
  geom_path(data=wildboar_closeup, aes(x=E, y=N, color=time0)) +
  geom_point(data = s1, aes(x=E, y=N), colour="black", size=2) +
  geom_point(data = schreck_orientation, aes(x=ausrichung_meanE, y=ausrichung_meanN), 
             colour="brown", size=2) +
  geom_sf(data=poly)
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Adding closest active schreck (+ distance) to each observation
```{r }
### data
w<-wildschwein_BE  
sl<-schreck_locations

###
s<-data.frame(schreck_locations_ch) ## I had to transform it to a data frame, otherwise something was weird about the coordinates (geometry)
###

s1<-s[!duplicated(s$id),c("id","N","E")] # list with all Schrecks and their location


## reduce data to certain time frame
w1<-w %>% 
  filter (day!="Tag"&!is.na(day))
nrow(w1)

time<-Sys.time()
#for(j in 1:nrow(w1)){ # loop for each unique observation
for(j in 1:10000){ 
### look only at Schrecks that were active on that day
  s_on<-s[s$datum_on < w1[j,]$DatetimeUTC & s$datum_off > w1[j,]$DatetimeUTC,]
if(nrow(s_on)==0){w1[j,"closest_schreck"]<-"no_Schreck_on"}  else{     
 for(i in 1:nrow(s_on)){ ## calculate difference between current observation (j) and each schreck location
 s_on[i,"diff"]<-sqrt((w1[j,"N"]-s_on[i,"N"])^2+(w1[j,"E"]-s_on[i,"E"])^2)} ## add difference of current observation to location into file
#### look at the distance of the closest Schreck, only use it if less than 400m
  if(min(s_on$diff)>400){w1[j,"closest_schreck"]<-"no_Schreck_witin_400m"} else{
w1[j,"closest_schreck"]<-s_on[s_on$diff==min(s_on$diff),"id"]  ## add closest schreck to each wild boar location
w1[j,"distance_to_closest_schreck"]<-s_on[s_on$diff==min(s_on$diff),"diff"]
}}}
time-Sys.time()
## 

## crate new subset of s that has only one line per location (will be used during loop)

head(data.frame(w1))
table(w1$closest_schreck)

# If it's complete, let it run once and then save it